SU($n$) spin models near the Heisenberg critical point

### 2020-03-19 collective_sun/sun_perturbation.tex
started explaining the construction of operators in the 'shell basis'

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$ (with respect to $H_0$).  This projector essentially
picks off the part of $X_{\m w}$ that is strictly off-diagonal with
respect to the ground- and excited-state manifolds $\E_0$ and
$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
states of definite excitation energy when acting on PS states $\M_0$,
we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ of the
interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
is
\begin{align}
  \p{\m h \circ_a \m w}_k
  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
\end{align}
where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
explicitly exclude terms with $p\in k$ that vanish due to the fact
that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
dimension-$M$ tensor $\tilde{\m h}$ with entries
\begin{align}
  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
\end{align}
for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
entries $\tilde h_k w_k$.  These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\check{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
symmetries of $H_0$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem.  We
discuss this reduction in Appendix \ref{sec:symmetries}.

Defining $\m w_\Delta$ as the projection of the tensor $\m w$ onto the
$\Delta$-eigenspace of $\check{\m h}$, we can expand
\begin{align}
  \m w = \sum_\Delta \m w_\Delta,
  &&
  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
\end{align}
and in turn
\begin{align}
  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
  &&
  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
\end{align}
Substituting these expansions into \eqref{eq:diagnosis}, we find
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
\end{align}
which implies that the action of $X_{\m w_\Delta}$ on a PS state
$\ket\psi\in\M_0$ generates a state of definite interaction energy
$E_0+\Delta$.  It follows that
$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
  w_\Delta} \P_0$, and so
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
\end{align}
The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
includes $r$-body operators for all positive integers $r\le2M$.  We
provide a general prescription for simplifying such products in
Appendix \ref{sec:operator_product}, and explicitly work out the
single-body ($M=1$) and two-body ($M=2$) cases in Appendices
\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
single-body operator ($M=1$), then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
  \p{\col{X}^2 - N \col{X^2}},
  &&
  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
\end{align}
which in the SU(2) case implies that an inhomogeneous magnetic field
($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
perturbation theory.

@@

In order to project operators onto $\ul{\M}_{r_{\t{max}}}$, we need to
construct a basis for each $\M_r$ with $r\le r_{\t{max}}$.
\red{[TODO: finish section]}

@@

  \equiv \p{\sum_a \m h \circ_a \m w - \tilde{\m h} * \m w}.

@@

  \label{eq:multi_body_eig}

@@

Solving the eigenvalue problem in \eqref{eq:multi_body_eig} to find
$M$-body perturbations $X_{\m w}$ that generate states of definite
interaction energy nominally requires diagonalizing the matrix
$\hat{\m h}$ with dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
interaction Hamiltonian $H_0$ has any spatial symmetries, however,
then we can use these symmetries to reduce the complexity of this
eigenvalue problem.  Translational invariance, for example, can reduce
this problem to that of diagonalizing a matrix with dimensions
$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
and isotropy can further reduce the problem size by factors that
depend on the lattice geometry.  The complexity of diagonalizing a
matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
of $d$ by a factor of 2 yields practically significant gains (a factor
of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
general prescription for exploiting the spatial symmetries of $H_0$ to
simplify the eigenvalue problem in \eqref{eq:multi_body_eig}.

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_eig} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set of
choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
  as well as the dependence of $\S_M$ on any other factors such as
  lattice geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of a translationally invariant and
isotropic system with two-body perturbations, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.

For simplicity, we first assume that the $M$-body operator $X_{\m w}$
obeys the same spatial symmetries as $H_0$.  For any equivalence class
$s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all
choices $k\in s$, as well as the projector $\P_\S$ onto the space of
uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.  We then expand the
coefficient vector $\v{\m w}$ in this symmetrized basis:
\begin{align}
  \v{\m w} = \sum_{k\in\C_N\p{M}} w_k \ket{k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
projection of $\hat{\m h}$ onto the symmetric subspace:
\begin{align}
  \hat{\m h}_{\S_M}
  \equiv \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \ket{s} \sp{
    \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
    \f{h_{p k_a}}{\sqrt{\abs{s}\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}}
    - \tilde h_s \bra{s}},
\end{align}
where $\tilde h_s\equiv \tilde h_k$ for any $k\in s$; $k_a$ is the
$a$-th element of $k$;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,
but with the $a$-th number $k_a$ replaced by $p$; and $\sp{k_{a:p}}$
is the equivalence class of $k_{a:p}$.  The reduced multi-body
eigenvalue problem is then
\begin{align}
  \hat{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
\end{align}
More generally, even if the couplings $\v{\m w}$ of $X_{\m w}$ do not
obey the same spatial symmetries as $H_0$, these symmetries generally
endow $\check{\m h}$ with a block-diagonal structure. \red{[TODO:
  finish discussion/derivation]}

@@

The process of constructing and simplifying diagrams to compute the
coefficients $w\p{g}$ of the diagrammatic operator product expansion
in \eqref{eq:sym_prod_group} is systematic enough to be converted into
an algorithm for execution on a computer.  The last step necessary to
automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
the multi-spin operators $X\p{g}$ in a basis for the permutationally
symmetric manifold $\M_0$.  Without loss of generality, we consider a
single permutationally symmetric $M$-spin operator $X$; our task is
essentially to find coefficients of the expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b\in\B_0} \bk{a|X|b} \op{a}{b},
\end{align}
where $\set{\ket{a}:a\in\B_0}$ is complete bases for $\M_0$, and the
choice of $M$ spins on which $X$ acts is arbitrary.  In a system of
$N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
labeled by the occupation number of each single-spin state, namely
$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_\mu\ge 0$ and
$\sum_\mu a_\mu=N$.  Written out explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
  &&
  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of
distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
$N$ spins to $n$ states by $\A_n\p{N}$, such that the basis
$\set{\ket{a}:a\in\A_n\p{N}}$ spans permutationally symmetric manifold
of $N$ spins.  Identifying this basis allows us to expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are
evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
$\alpha$ and $\beta$ above merely to keep the expression symmetric
with respect to transposition; in practice, one can simply sum over
$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
abuse of notation, the operator $X$ on the left of
\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
(out of $N$), whereas the operator $X$ on the right of
\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
states $\ket\alpha,\ket\beta$.

>>>>

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$ (with respect to $H_0$).  This projector essentially
picks off the part of $X_{\m w}$ that is strictly off-diagonal with
respect to the ground- and excited-state manifolds $\E_0$ and
$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
states of definite excitation energy when acting on PS states $\M_0$,
we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ of the
interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
is
\begin{align}
  \p{\m h \circ_a \m w}_k
  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
\end{align}
where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
explicitly exclude terms with $p\in k$ that vanish due to the fact
that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
dimension-$M$ tensor $\tilde{\m h}$ with entries
\begin{align}
  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
\end{align}
for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
entries $\tilde h_k w_k$.  These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
  \label{eq:multi_body_eig_op}
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\hat{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
symmetries of $H_0$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem.  We
discuss this reduction in Appendix \ref{sec:symmetries}.

Defining $\m w_\Delta$ as the projection of the tensor $\m w$ onto the
$\Delta$-eigenspace of $\hat{\m h}$, we can expand
\begin{align}
  \m w = \sum_\Delta \m w_\Delta,
  &&
  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
\end{align}
and in turn
\begin{align}
  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
  &&
  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
\end{align}
Substituting these expansions into \eqref{eq:diagnosis}, we find
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
\end{align}
which implies that the action of $X_{\m w_\Delta}$ on a PS state
$\ket\psi\in\M_0$ generates a state of definite interaction energy
$E_0+\Delta$.  It follows that
$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
  w_\Delta} \P_0$, and so
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
\end{align}
The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
includes $r$-body operators for all positive integers $r\le2M$.  We
provide a general prescription for simplifying such products in
Appendix \ref{sec:operator_product}, and explicitly work out the
single-body ($M=1$) and two-body ($M=2$) cases in Appendices
\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
single-body operator ($M=1$), then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
  \p{\col{X}^2 - N \col{X^2}},
  &&
  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
\end{align}
which in the SU(2) case implies that an inhomogeneous magnetic field
($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
perturbation theory.

@@

In order to project an operator $X_{\m w}$ onto
$\ul{\M}_{r_{\t{max}}}$, we first need to construct a basis for
$\ul{\M}_{r_{\t{max}}}$, and then compute all matrix elements of
$X_{\m w}$ in the constructed basis.  Our strategy will be to first
identify a basis for $\M_0$, and then find, for each
$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
$\ul{\M}_{r_{\t{max}}}$ then take the form
\begin{align}
  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
\end{align}
In order to compute these matrix elements, we simplify products of the
form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
$\M_0$, in Appendix \ref{sec:operator_product}.

A suitable basis for $\M_0$ is the set of PS states
$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
$m_\mu$ of the single-spin state $\mu$.

\red{[TODO: finish section]}

@@

  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.

@@

  \label{eq:multi_body_eig_mat}

@@

Solving the eigenvalue problem in \eqref{eq:multi_body_eig_mat} to
find $M$-body perturbations $X_{\m w}$ that generate states of
definite interaction energy nominally requires diagonalizing the
matrix $\hat{\m h}$ with dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
interaction Hamiltonian $H_0$ has any spatial symmetries, however,
then we can use these symmetries to reduce the complexity of this
eigenvalue problem.  Translational invariance, for example, can reduce
this problem to that of diagonalizing a matrix with dimensions
$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
and isotropy can further reduce the problem size by factors that
depend on the lattice geometry.  The complexity of diagonalizing a
matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
of $d$ by a factor of 2 yields practically significant gains (a factor
of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
general prescription for exploiting the spatial symmetries of $H_0$ to
simplify the eigenvalue problem in \eqref{eq:multi_body_eig_mat}.

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_eig_mat} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set of
choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
  as well as the dependence of $\S_M$ on any other factors such as
  lattice geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of a translationally invariant and
isotropic system with two-body perturbations, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.

For simplicity, we first assume that the $M$-body operator $X_{\m w}$
obeys the same spatial symmetries as $H_0$.  For any equivalence class
$s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all
choices $k\in s$, as well as the projector $\P_\S$ onto the space of
uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.  We then expand the
coefficient vector $\v{\m w}$ in this symmetrized basis:
\begin{align}
  \v{\m w} = \sum_{k\in\C_N\p{M}} w_k \ket{k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
projection of $\hat{\m h}$ onto the symmetric subspace:
\begin{align}
  \hat{\m h}_{\S_M}
  \equiv \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \ket{s} \sp{
    \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
    \f{h_{p k_a}}{\sqrt{\abs{s}\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}}
    - \tilde h_s \bra{s}},
\end{align}
where $\tilde h_s\equiv \tilde h_k$ for any $k\in s$; $k_a$ is the
$a$-th element of $k$;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,
but with the $a$-th number $k_a$ replaced by $p$; and $\sp{k_{a:p}}$
is the equivalence class of $k_{a:p}$.  The reduced multi-body
eigenvalue problem is then
\begin{align}
  \hat{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
\end{align}
More generally, even if the couplings $\v{\m w}$ of $X_{\m w}$ do not
obey the spatial symmetries as $H_0$, these symmetries generally endow
$\hat{\m h}$ with a block-diagonal structure. \red{[TODO: finish
  discussion/derivation]}

@@

The process of constructing and simplifying diagrams to compute the
coefficients $w\p{g}$ of the diagrammatic operator product expansion
in \eqref{eq:sym_prod_group} is systematic enough to be converted into
an algorithm for execution on a computer.  The last step necessary to
automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
the multi-spin operators $X\p{g}$ in a basis for the permutationally
symmetric manifold $\M_0$.  Without loss of generality, we consider a
single permutationally symmetric $M$-spin operator $X$; our task is
essentially to find coefficients of the expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b\in\B_0} \bk{a|X|b} \op{a}{b},
\end{align}
where $\set{\ket{a}:a\in\B_0}$ is complete bases for $\M_0$, and the
choice of $M$ spins on which $X$ acts is arbitrary.  In a system of
$N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
labeled by the occupation number of each single-spin state, namely
$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_\mu\ge 0$ and
$\sum_\mu a_\mu=N$.  Written out explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
  &&
  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of
distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
$N$ (identical) spins to $n$ (distinguishable) states by $\A_n\p{N}$,
such that the basis $\set{\ket{a}:a\in\A_n\p{N}}$ spans
permutationally symmetric manifold of $N$ spins.  Identifying this
basis allows us to expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are
evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
$\alpha$ and $\beta$ above merely to keep the expression symmetric
with respect to transposition; in practice, one can simply sum over
$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
abuse of notation, the operator $X$ on the left of
\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
(out of $N$), whereas the operator $X$ on the right of
\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
states $\ket\alpha,\ket\beta$.
